\clearpage
\section{Exact Algorithms}\label{sec:solving:exact}

\subsection{DHEA Solver}
\label{sec:exact:dhea}
The DHEA solver, defined and implemented
by \citet{ljubic2005solving}
(more concisely described in \citet{ljubic2006algorithmic}) finds exact solutions to the PCSTP
by first applying
reduction tests to the input graph (See Section \ref{sec:solving:pre}) and
then transforming the resulting problem into related problem of finding a minimum
aborescence in a directed graph. It is the first example of a procedure designed for
finding provably optimal solutions for the PCSTP.

The resulting SAP is then solved by branch-and-cut using a separation procedure
similar to the one defined in \citet{lucena2004strong}
(Section \ref{sec:lower:gsec}) and a new primal heuristics for
finding good incumbents at every node.

The solver has as a dependency one of the MIP solvers CPLEX or LEDA.
 \paragraph{Reducing the PCSTP to PCSAP}
 Citing stronger LP relaxations, \citet{ljubic2005solving} reduces an input PCSTP instance
 into a related problem in a directed graph which essentially is a constrained version
 of a Prize-Collecting Steiner Aborescence Problem (PCSAP).

 Given an instance of the PCSTP, with graph $G = (V, E, c, p)$ and terminals $N \subset V$,
 we define an instance
 of the PCSAP with directed graph, $G_A = (V_A, A, c')$, where $c'$ define the arc costs,
 by first adding an
 artificial root vector, $r$, to $V$
 $$V_A = \{r\} \cup V\mathnormal{.}$$
 Then we define the arc set by adding bi-directional arcs for each edge in $E$ and
 adding an arc from $r$ to each terminal,
 $$A = \{(i,j) \mid (i,j) \in E \} \cup \{ (j,i) \mid (i,j) \in E \} \cup \{(r, i) \mid i \in N\}\mathnormal{.}$$
 We define arc-costs by subtracting the prize of the arc's endpoint from the original cost of the
 corresponding edge in $G$. Arcs from the root have cost equal to the negative prize of the endpoint vertex. This gives us,
 $$c'_{ij} =
 \begin{cases}
   -p_j & i = r \\
   c_{ij} - p_j & \text{otherwise.}
 \end{cases}
 $$
Figure (\ref{fig:pcstptosap}) shows an example of this kind of reduction.
 \begin{figure}[h]\centering
  \begin{subfigure}[t]{0.47\linewidth}
    
    \begin{tikzpicture}[auto, node distance=1.5 cm]
      %Nodes
  \node[terminal, label={12}] (a) {a};
  \node[steiner] (b) [right=of a] {b};
  \node[steiner] (c) [right=of b] {c};
  \node[terminal, label={10}] (d) [above =of c] {d};
  \node[steiner] (e) [left=of d] {e};
  \node[steiner] (f) [left=of e] {f};
  \node[terminal, label={3}] (g) [above right=0.75 and 1.3 of c] {g};
  % Edges
  \begin{scope}[every edge/.style={draw=black, thick}]
    \draw (a) edge node[below]{4} (b);
    \draw (b) edge node[near start]{5} (d);
    \draw (b) edge node[below]{8} (c);
    \draw (c) edge node{3} (d);
    \draw (c) edge node{2} (g);
    \draw (c) edge node[near start]{5} (e);
    \draw (d) edge node{6} (e);
    \draw (d) edge node{10} (g);
    \draw (e) edge node{1} (f);
  \end{scope}
    \end{tikzpicture}
    \caption{PCSTP Instance (\ref{fig:pcstp:01})}
  \end{subfigure}
  \begin{subfigure}[t]{0.47\linewidth}
    \begin{tikzpicture}[auto, node distance=1.7 cm]
      % Nodes
      \node[terminal] (a) {a};
      \node[steiner] (b) [right=of a] {b};
      \node[steiner] (c) [right=of b] {c};
      \node[terminal] (d) [above =of c] {d};
      \node[steiner] (e) [left=of d] {e};
      \node[steiner] (f) [left=of e] {f};
      \node[terminal] (g) [above right=0.85 and 1.3 of c] {g};
      \node[terminal, root] (r) [below= of g] {r};
      % Edges
      \begin{scope}[every edge/.style={draw=black, thick},
        every node/.style={circle, inner sep=.7}]
        \draw[->] (a) edge[bend right=10] node[below]{4} (b);
        \draw[<-] (a) edge[bend left=10] node{-8} (b);
        \draw[->] (b) edge[bend right=10] node[near start,swap] {-5} (d);
        \draw[<-] (b) edge[bend left=10] node[near start] {5} (d);
        \draw[<->] (b) edge node[below]{8} (c);
        \draw[->] (c) edge[bend right=10] node[swap]{-7} (d);
        \draw[<-] (c) edge[bend left=10] node{3} (d);
        \draw[<-] (c) edge[bend left=10] node{2} (g);
        \draw[->] (c) edge[bend right=10] node[swap]{-1} (g);
        \draw[<->] (c) edge node[near end]{5} (e);
        \draw[->] (d) edge[bend right=10, swap] node{6} (e);
        \draw[<-] (d) edge[bend left=10] node{-4} (e);
        \draw[->] (d) edge[bend right=10] node[swap]{7} (g);
        \draw[<-] (d) edge[bend left=10] node{0} (g);
        \draw[<->] (e) edge node{1} (f);
        \draw[->] (r) edge[bend right=90, looseness=2] node{-10} (d);
        \draw[->] (r) edge node{-3} (g);
        \draw[->] (r) edge[bend left=40] node{-12} (a);
      \end{scope}
    \end{tikzpicture}
    \caption{Instance (\ref{fig:pcstp:01}) as a PCSAP instance. Root node is coloured blue.}
  \end{subfigure}
  \caption{Reduction from PCSTP to SAP.}
  \label{fig:pcstptosap}
\end{figure}

Clearly, if we
limit the out-degree of the root vector, $r$, to $\delta^+(r) \leq 1$ in any feasible
aborescence $SA$, then finding the aborescence in $G_A$ which minimises the function
$$c(SA) = \sum_{a \in A_{SA}} c'_{a} + \sum_{v \in N} p_v$$
is equivalent to finding a minimal prize-collecting Steiner tree in $G$.

\paragraph{ILP Formulation}

Finding such a minimum ``prize-collecting'' aborescence is modelled by \citet{ljubic2005solving}
as the ILP in Fomulation (\ref{form:exact:cut}).

 \begin{formulation}[h!]
   \begin{subequations}
     \begin{alignat}{3} 
       &\underset{x,y}{\text{minimize}}
       & & \sum_{ij \in A} c'_{ij} x_{ij} +  \sum_{v \in V} p_v  & \\
       & \text{subject to}\quad
       & & \sum_{j \in A} x_{rj} = 1 &&  \label{form:exact:pcsap:root}\\
       &&& \sum_{ji \in A} x_{ji} = y_i  && \forall i \in V_A \label{form:exact:pcsap:prize}\\
       &&& x(\delta^-(S)) \geq y_k && \forall k \in S, S \subseteq V_A \setminus \{r\}
       \label{form:exact:pcsap:conn}\\
       \intertext{ILP:}
       &&& x_a \in \BB  && \forall a \in A \\
       &&& y_v \in \BB  && \forall v \in V_A \\
       \intertext{LP:}
       &&& 0 \leq x_a \leq 1  && \forall a \in A \\
       &&& 0 \leq y_v \leq 1  && \forall v \in V_A \\
       \intertext{To ensure bijection:}
       &&& x_{rj} \leq 1 - y_i && \forall i,j \in V_A \setminus \{r\}; i < j
       \label{form:exact:bijection}\\
       \intertext{Nonterminal Degree:}
       &&& \sum_{j \in A}x_{ij} \leq \sum_{j \in A}x_{ji}
       && \forall i \in V_A \setminus (N \cup \{r\})
       \label{form:exact:strength}\\
       \intertext{Single Orientation:}
       &&& x_{ij} + x_{ji} \leq y_i  && \forall i \in V_A \setminus \{r\}; \forall (i,j) \in A
       \label{form:exact:orientation}
     \end{alignat}\label{form:exact:cut}
   \end{subequations}
   \caption{(CUT-IP) GSEC formulation of the constrained PCSAP.}
 \end{formulation}

 As usual, the ILP formulation has two decision vectors, $x$ and $y$, where the former denotes arcs
 in the solution arborescence and the latter which vertices are in the arborescence. Constraints-wise,
 we have
 constraint (\ref{form:exact:pcsap:root}), which ensures that exactly one arc begins in the artificial
 root in any feasible aborescence. Constraint (\ref{form:exact:pcsap:prize}) ensures
 that a vertex can be selected iff it has an incoming arc.
 Constraints (\ref{form:exact:pcsap:conn}), called \textit{connectivity constraints},
 ensures the connectivity of feasible aborescences by requiring that
 each vertex subset containing a selected vertex must have an incoming arc.
 Together, these constraints ensure that any feasible solution is an aborescence rooted
 in $r$ with $\delta^+(r) = 1$.

 It is worth noting that for a given subset $S \subseteq V$, adding constraints of type
 (\ref{form:exact:pcsap:prize}) for each $v \in S$ to a corresponding connectivity constraint
  gives us generalised subtour elimination constraints:
 \begin{align*}
   -x(\delta^-(S)) &\leq -y_k \\
   \sum_{i \in S} \sum_{ji \in A} x_{ji} - x (\delta^-(S)) &\leq \sum_{v \in S} - y_k \\
   \sum_{i \in S} \sum_{j \in S} x_{ij} &\leq \sum_{v \in S} - y_k \mathnormal{.}
 \end{align*}

 Adding only constraints (\ref{form:exact:pcsap:root})-(\ref{form:exact:pcsap:conn})
 has the unfortunate consequence of having multiple 
 PCSAP solutions
 correspond
 to a single PCSTP solution. To achieve a bijection between the PCSAP solutions and the
 PCSTP solutions, constraint (\ref{form:exact:bijection}) allows only the chosen vertex
 with the lowest index to be adjacent to the root vertex.
 
 Additionally, \citet{ljubic2005solving} observe that each nonterminal in an optimal solution
 \textit{must not} be a leaf in the arborescence. Hence, they add constraint
 (\ref{form:exact:strength}) to strengthen the formulation. And finally, they
 also capture the fact that there cannot be anti-parallel in a feasible aborescence in
 constraint (\ref{form:exact:orientation}). While this constraint is captured implicitly
 in the previous constraints and adding them increases the size of the formulation,
 \citet{ljubic2005solving} report improved running times for adding them as it avoids
 having to implicitly separate the in the separation procedure.
 
\paragraph{Branch \& Bound}

For finding lower bounds, the LP relaxation of (\ref{form:exact:cut}) is solved at each node.
Constraints of type (\ref{form:exact:pcsap:conn}) are separated through solving the max flow (min cut)
problem on a support graph (similar to the procedure in Section \ref{sec:lower:gsec}). 
The max flow problem is solved for every terminal $v \in N$ and returns both the minimum cut closest to
the root, and the minimum cut $(S,T)$ ``closest'' to the sink with $v \in T$.
Violated constraints are iteratively found by increasing the
capacity of all edges crossing a min cut to $1$ and rerunning the max flow procedure.

Primal heuristics based on the solution of LP relaxations are used at each node in the branch and bound
tree to discover improved incumbents.

The rest of the branch and bound procedure is handled by an underlying MIP solver (CPLEX or LEDA).
\subsection{SCIP-Jack Solver}
\label{sec:exact:scipj}
The SCIP-Jack solver is a set of plugins for the SCIP general purpose optimisation suite capable
of generating exact solutions to a variety of STP variants in graphs \cite{gamrath2017scip}.

As with the DHEA solver (Section \ref{sec:exact:dhea}), SCIP-Jack solves the PCSTP by first applying
reduction tests (Section \ref{sec:solving:pre}) to the input graph before transforming problem
instances to a constrained version of the SAP and solving it with branch and cut.
The main difference here is in that the DHEA solver works with a prize-collecting variant
 of the SAP problem whereas SCIP-Jack works with the regular version of the SAP.

\subsubsection{Reduction to SAP}

$$N_A = \{v' \mid v \in N\}$$
$$V_A = V \cup \{r\} \cup N_A$$
$$c`_{ij} =
\begin{cases}
  c_{ij} & (i,j) \in E \\
  p_v & i = r, j \in N_A , j = v' \\
  0 & i = r, j \in N \\
  0 & i \in N, j \in N_A
\end{cases}\textnormal{.}$$
\begin{align*}
A = &\{(i,j) \mid (i,j) \in E \} \cup \{ (j,i) \mid (i,j) \in E \} \:\cup \\
&\{(r, i) \mid i \in N\} \cup \{(r,i) \mid i \in N_A\} \cup \{(i,j) \mid i \in N, j \in N_A\}\mathnormal{.}
\end{align*}

\begin{figure}[h]\centering
  \begin{subfigure}[t]{0.47\linewidth}
    
    \begin{tikzpicture}[auto, node distance=1.5 cm]
      %Nodes
  \node[terminal, label={12}] (a) {a};
  \node[steiner] (b) [right=of a] {b};
  \node[steiner] (c) [right=of b] {c};
  \node[terminal, label={10}] (d) [above =of c] {d};
  \node[steiner] (e) [left=of d] {e};
  \node[steiner] (f) [left=of e] {f};
  \node[terminal, label={3}] (g) [above right=0.75 and 1.3 of c] {g};
  % Edges
  \begin{scope}[every edge/.style={draw=black, thick}]
    \draw (a) edge node[below]{4} (b);
    \draw (b) edge node[near start]{5} (d);
    \draw (b) edge node[below]{8} (c);
    \draw (c) edge node{3} (d);
    \draw (c) edge node{2} (g);
    \draw (c) edge node[near start]{5} (e);
    \draw (d) edge node{6} (e);
    \draw (d) edge node{10} (g);
    \draw (e) edge node{1} (f);
  \end{scope}
    \end{tikzpicture}
    \caption{PCSTP Instance (\ref{fig:pcstp:01})}
  \end{subfigure}
  \begin{subfigure}[t]{0.47\linewidth}
    \begin{tikzpicture}[auto, node distance=1.7 cm]
      % Nodes
      \node[steiner] (a) {a};
      \node[terminal] (aa) [below=1cm of a] {a'};
      \node[steiner] (b) [right=of a] {b};
      \node[steiner] (c) [right=of b] {c};
      \node[steiner] (d) [above =of c] {d};
      \node[terminal] (dd) [above= 0.8cm of d] {d'};
      \node[steiner] (e) [left=of d] {e};
      \node[steiner] (f) [left=of e] {f};
      \node[steiner] (g) [above right=0.85 and 1.3 of c] {g};
      \node[terminal] (gg) [below= 0.8cm of g] {g'};
      \node[terminal, root] (r) [right= of g] {r};
      % Edges
      \begin{scope}[every edge/.style={draw=black, thick},
        every node/.style={circle, inner sep=.7}]
        \draw[<->] (a) edge node[below]{4} (b);
        \draw[->] (a) edge node {0} (aa);
        \draw[<->] (b) edge node[near start] {5} (d);
        \draw[<->] (b) edge node[below]{8} (c);
        \draw[<->] (c) edge node{3} (d);
        \draw[<->] (c) edge node{2} (g);
        \draw[<->] (c) edge node[near end]{5} (e);
        \draw[->] (d) edge node {0} (dd);
        \draw[<->] (d) edge node{6} (e);
        \draw[<->] (d) edge node[swap]{7} (g);
        \draw[<->] (e) edge node{1} (f);
        \draw[->] (g) edge node{0} (gg);
        \draw[->] (r) edge[bend right=40] node{0} (d);
        \draw[->] (r) edge[bend right=40] node{10} (dd);
        \draw[->] (r) edge node{0} (g);
        \draw[->] (r) edge node{3} (gg);
        \draw[->] (r) edge[bend left=45, looseness=1.0] node{0} (a);
        \draw[->] (r) edge[bend left=50, looseness=0.7] node{12} (aa);
      \end{scope}
    \end{tikzpicture}
    \caption{Instance (\ref{fig:pcstp:01}) as a SAP instance. Root node is coloured blue.}
  \end{subfigure}
  \caption{Reduction from PCSTP to SAP.}
  \label{fig:pcstptosap}
\end{figure}


 \subsubsection{ILP Formulation}

 \begin{formulation}[h!]
   \begin{subequations}
     \begin{alignat}{3} 
       &\underset{x}{\text{minimize}}
       & & \sum_{ij \in A} c'_{ij} x_{ij} & \\
       & \text{subject to}\quad
       && \sum_{v \in \delta^+(r), c_{rv} = 0} y_{rv} && 
       \label{form:sap:root}\\
       &&& x(\delta^+(S)) \geq 1 && \forall S \subset V, r \in S, N \not\subset S
       \label{form:exact:sap:cut}\\
       &&& x(\delta^-(v))
       \begin{cases}
         = 0 & v = r \\
         = 1 & v \in N \\
         \leq 1 & v \in V_A \setminus (N \cup \{r\})
       \end{cases}
       && \forall v \in V_A \label{form:exact:sap:indeg}\\
       &&& x(\delta^-(v)) \leq x(\delta^+(v)) && \forall v \in V \setminus N
       \label{form:exact:sap:deg}\\
       &&& x(\delta^-(v)) \geq x_{vu} && \forall v \in V \setminus N, \forall u \in \delta^+(v)
       \label{form:exact:sap:conn}\\
       \intertext{ILP:}
       &&& x_a \in \BB  && \forall a \in A \\
       \intertext{LP:}
       &&& 0 \leq x_a \leq 1  && \forall a \in A \\
       \intertext{To ensure bijection:}
       &&& \sum_{i \in \delta^-(j)} x_{ij} + x_{rv}
       && \forall v = 1,...,|V_A|, j = 1,...,i-1
       \label{form:sap:bijection}
     \end{alignat}\label{form:sap:flow}
   \end{subequations}
   \caption{(FLW-IP) Flow-balance directed cut formulation of the constrained SAP.}
 \end{formulation}


%%% Local Variables:
%%% TeX-master: "report"
%%% reftex-default-bibliography: ("lit.bib")
%%% End:
