\section{Heuristics}\label{sec:solving:heuristics}
There has been a number of heuristics defined for the \gls{pcstp}:

\begin{itemize}
\item \citet{canuto2001local} presented a local search based heuristics which we
   will detail below.
 \item \citet{Ljubic:2004:memetic} presented a heuristics with inspiration in genetic
   algorithms.
 \item \citet{ljubic2005solving} presented a LP relation based heuristic for their branch
   and bound algorithm which we detail below.
\item \citet{fu2014knowledge} presented a tabu search heuristic.
 \item \citet{akhmedov2016divide} presented a divide and conquer algorithm based on clustering
    for applications within bioinformatics.
\end{itemize}
We have chosen to show in this section, one search based and one LP based heuristic to give
an overview of general heuristics methods can be applied to the \gls{pcstp}.
\subsection{Local Search with Perturbation and Relinking}\label{sec:canuto-search}

\citet{canuto2001local} presented a multi-phase heuristics which makes use of the GW Algorithm.
 Algorithm \ref{alg:heuristics:canuto} sketches the full procedure. 

 The heuristics can be seen as a multi start heuristics which uses the GW Algorithm
 to generate good initial
 solutions (line \ref{alg:canuto:line:gw})
 which it progressively refines using a hill climbing local search
 (line \ref{alg:canuto:line:hc}),
 and then a path relinking scheme against a previous good solution
 (line \ref{alg:canuto:line:relink}).
 Perturbation of the prize vector, $p$, is used to push the GW Algorithm to generate
 different solutions on multiple starts, and finally, a
 variable neighbourhood search is applied as a
 post processing step to refine the best candidate solution found.

 \begin{algorithm}[h!]
   \begin{algorithmic}[1]
     \Input{Graph $G = (V, E, c, p)$.}
     \Output{Tree $\tilde{T} \subseteq G$.}
     \Procedure{Canuto-LocalSearch}{}
     \State $E \gets \emptyset$, $\hat{p} \gets p$, $\hat{z}^* \gets \infty$
     \For{$i \in 1...\mathit{max\_iterations}$}
     \State $S \gets \Call{GW-Algorithm}{V,E, c, \hat{p}}$ \label{alg:canuto:line:gw}
     \If{$S$ hasn't been seen in previous iterations}
     \State $S' \gets \Call{Hill-Climbing}{S', G = (V, E, c, p)}$ \label{alg:canuto:line:hc}
     \State Add $S'$ to $E$ if it satisfies the necessary criteria \label{alg:canuto:line:elite}
     \State Pick random $Y \in E$
     \State $S' \gets \Call{Relink}{S', Y}$ \label{alg:canuto:line:relink}
     \If{$c(S') < \hat{z}^*$}
     \State $\hat{z}^* \gets c(S')$
     \State $S^* \gets S'$
     \EndIf
     \EndIf
     \State Perturbate $\hat{p}$
     \EndFor
     \State $S^* \gets \Call{VNS}{S^*, G = (V, E, c, p)}$\label{alg:canuto:line:vns}
     \State \Return $\Call{MST}{G[S^*]}$
     \EndProcedure
 \end{algorithmic}
 \caption{The heuristics defined by \citet{canuto2001local}.}\label{alg:heuristics:canuto}
 \end{algorithm}

 In this section, we
 describe further the characteristics of the local search, the
 path relinking, and the postprocessing used in the algorithm.
\paragraph{Local Search with Perturbations}
\citet{canuto2001local} define their search space as the space of minimum spanning
trees on all graphs induced by a subset of vertices
 $S \subset V$. If $G[S]$ is the subgraph induced by the vertex subset $S \subseteq V$, then
 this search space can formally be defined as
\[\mathcal{S_G} = \{MST(G[S]) : S \subseteq V \}\mathnormal{.}\]
A neighbour to a given MST is then any MST of a subgraph induced by a subset which differs from $S$
 by exactly one vertex, that is,
\[\mathcal{N}(S) = \Big\{MST(G[S']) : S' \subseteq V \Bigm| |S' \triangle S| = 1 \Big\}\mathnormal{.}\] 
Clearly this neighbourhood structure means that all feasible solutions are reachable
given any initial subset as starting point. Additionally, they define the
$k$'th order neighbourhood as,
\[\mathcal{N}^k(S) = \Big\{MST(G[S']) : S' \subseteq V \Bigm| |S' \triangle S| = k \Big\}\mathnormal{.}\] 

To reduce clutter we denote the cost
of a given subset, $S \subseteq V$, of vertices as the cost
of the MST of the graph induced by $S$, that is
\[c(S) = c(MST(G[S]))\mathnormal{.}\]

The search algorithm itself is just a plain hill-climbing algorithm which repeatedly
 moves to the first
 found neighbouring solution with lower cost until it arrives at a local minima.

 To overcome the problem of running into the same local minima repeatedly,
 \citet{canuto2001local} employ two perturbation schemes which are alternately applied
 on even and odds iterations:
 \begin{enumerate}
 \item A fixed percentage of vertices which have appear both in the initial solution generated
   by the GW Algorithm \textit{and}  in the solution subsequently generated by
    the hill climbing procedure have their prizes
    set to zero.
  \item A perturbation factor, $\alpha$, is randomly picked within a fixed interval $[1 -a, 1+a]$
    and prizes are set to $p_v \gets \alpha p_v$ for all $v \in V$.
 \end{enumerate}

\paragraph{Path Relinking}
For every iteration of the main loop, the relinking procedure in Algorithm \ref{heuristics:canuto:relink}
is applied to a solution generated by the hill climbing procedure and 
a random \textit{elite} solution which is picked from the
persistent pool of elite solutions, $E$.

The relinking procedure basically transforms $S$ into $Y$ one vertex at a time in a greedy fashion.
The solution along the way which has the best objective value is then returned.
 \begin{algorithm}[h!]
   \begin{algorithmic}[1]
     \Procedure{Apply}{S, v}
     \If{$v \in S$}
     \State \Return $S \setminus \{v\}$
     \Else
     \State \Return $S \cup \{v\}$
     \EndIf
     \EndProcedure
     \Procedure{Relink}{S, Y}
     \State $D \gets S \triangle Y$
     \State $S' \gets S$
     \State $\tilde{S} \gets S$
     \While{$S' \neq Y$}
     \State $v^* \gets \argmin_{v \in D} c(\Call{Apply}{S,v})$
     \State $S' \gets \Call{Apply}{S', v^*}$
     \State if $c(S') < c(\tilde{S})$ then set $\tilde{S} \gets S'$
     \State $D \gets D \setminus \{v\}$
     \EndWhile
     \State \Return $\tilde{S}$
     \EndProcedure
 \end{algorithmic}
 \caption{The relinking scheme used by \citet{canuto2001local}.}\label{heuristics:canuto:relink}
 \end{algorithm}

 Entry into $E$ for a subset $S$ is determined by two criteria:
 either we must
 \[c(\tilde{S}) < \min_{S \in E} c(S)\mathnormal{,}\]
 or we must have both
 \[c(\tilde{S}) < \max_{S \in E} c(S)\]
  and that the Hamming distances between the characteristic vectors of $\tilde{S}$
   and \textit{all} solutions in $E$ are greater than $\rho |V|$ for some $\rho \leq 1$. 

\paragraph{Postprocessing}
Variable neighbourhood search (VNS) (See \citet{hansen2010variable}) is applied
to best candidate solution as a post-processing step. In principle, the VNS
procedure (shown in Algorithm \ref{alg:heuristics:canuto:vns}) is a hill climbing
search within a union of the first $k_{max}$ order neighbourhoods with an early stopping
 criteria.
\begin{algorithm}[h!]
   \begin{algorithmic}[1]
     \Procedure{VNS}{S}
     \For{$i \in 1...\mathit{max\_iterations}$}
     \State $k \gets 1$
     \While{$k  \leq k_{max}$}
     \State Pick random $S' \in \mathcal{N}^k(S)$
     \State $\tilde{S} \gets \Call{Hill-Climbing}{S'}$
     \If{$c(\tilde{S}) < c(S)$}
     \State $k \gets 1$
     \State $S \gets \tilde{S}$
     \Else
     \State $k \gets k + 1$
     \EndIf
     \EndWhile
     \EndFor
     \State \Return $S$
     \EndProcedure
 \end{algorithmic}
 \caption{The Variable Neighbourhood Search
   used by \citet{canuto2001local}.}\label{alg:heuristics:canuto:vns}
 \end{algorithm}

 As a postprocessing step, this procedure is only applied to the single best solution generated
  by all the previous steps.
\paragraph{Evaluation}
Based on the implementation of the GW Algorithm by \citet{Johnson:2000:PCS:338219.338637},
\citet{canuto2001local} performed experiments, detailing the objective function value
and running time of
\begin{itemize}
\item The GW Algorithm,
\item The multi-start local search procedure with perturbations
\item The above combined with path relinking, and
\item The full heuristics.
\end{itemize}

All three ``parts'' of the heuristics are shown to help improve upon the
objective value generated by the previous parts.
But in particular, \citeauthor{canuto2001local} report that adding the
path relinking scheme results in a stark increase
 in number of instances solved to optimality.

 \subsection{LP Based Heuristic}
 \label{sec:heuristics:lp}
 As part of their branch and bound algorithm for the \gls{pcstp}, \cite{ljubic2005solving} detail a primal
 heuristics which takes advantage of an optimal solution to the LP relaxation of the problem. While this
 is detailed for a directed version of the \gls{pcstp} (see Section~\ref{sec:exact:dhea}) we detail it
 here for the \gls{pcstp}.
 
 Given an optimal solution to the \gls{gsec} and LP relaxation of the ILP
 \[(\bd{ \bar x^*}, \bd{\bar y^*})\]
 of an instance of the \gls{pcstp} defined on the graph $G = (V, E, c, p)$,
 a \textit{good} feasible solution to the \gls{pcstp} as follows.

 Select an initial subset of vertices to be part of the solution based on their decision variable value,
 \[S = \{ i \in V \mid \bar y^*_i > 0.5 \}\mathnormal{.}\]
 \textit{Alternatively, use $\bar y^*_i$ as the probability for including $i$ in $S$.}
 We then augment $G$ to have costs relative to $\bd{\bar{x}}$. More specifically,
 we create the cost function
 \[c'_{ij} = 1 - \bar{x}^*_{ij}\]
 and let it replace $c$ in $G$ to give us the graph,
 \[G' = (V, E, c', p)\mathnormal{.}\]
 With $G'$ and $S$ in hand, we compute the \textit{distance network} of $S$ as
 \[G_S = (S, S \times S, d_S)\]
 where ${d_S}_{ij}$ is the length of the shortest path from $v_i$ to $v_j$ in $G'$.
 Since we use $G'$, a path is short if its edges have high values in the LP
 relaxation.

 Then we find a minimum spanning tree on $G_S$,
 \[T = (S, E_S, d_S) \mathnormal{.}\]
 Let $P_{ij}$ be the shortest path between vertices $i$ and $j$ in $G$, then
 we select a new --- larger --- subset of $V$ by selecting all vertices which
 are implicitly part of an edge in $T$ according to $d_S$, that is
 \[S' = \{ v \in V \mid \exists (i,j) \in E_S, v \in P_{ij}\}\mathnormal{.}\]

 Finally, let $G_H$ be the graph induced by $S'$ on $G$,
 \[G_H = (S', E_H, c, p)\]
 and a minimum spanning tree of $G_H$, $T' = (S', E_{T'}, c, p)$.

 The heuristic solution is then given by solving the \gls{pcstp}
 on $T'$. This can be done in linear time, e.g. by using the
 strong pruning procedure defined by
 \citet{Johnson:2000:PCS:338219.338637}
 (see Section~\ref{sec:approx:strongpruning}).

 
 
 
%%% Local Variables:
%%% TeX-master: "report"
%%% reftex-default-bibliography: ("lit.bib")
%%% End:
